Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle structure

### 2017-05-31 documents/TriangleScenarioArXiv/ts.tex
removed all discussion of the local markov property

    Now that the notion of a causal structure $\graph$ is defined along with some standard graph-theoretic notation, it is now possible to precisely define the notion of compatibility between $\graph$ and a probability distribution $\prob[\nodes]$ defined over the nodes of $\graph$. Intuitively, the absence of certain causal connections within $\graph$ should limit the set of possible observations that can be made on $\graph$. Specifically, the causal structure $\graph$ hypothesizes that each variable $n \in \nodes$ is causally influenced by its parents $\Pa[\graph]{n}$, then $n$ should be independent of all other variables, except its own descendants when conditioned over $\Pa[\graph]{n}$. This is known as the \term{local Markov property}:\footnote{Here $A \indep B \mid C$ denotes statistical independence between $A$ and $B$ conditional on $C$ where $A,B,C \subseteq \nodes$, i.e. for a given probability distribution $\prob[\nodes]$, $\prob[AB \mid C] = \prob[A \mid C]\prob[B \mid C]$.}
    \[ n \indep \br{\nodes \setminus \De[\graph]{n}} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:local_markov_property}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. Alternatively yet equivalently, the notion of compatibility can be interpreted as follows: a distribution $\prob[\nodes]$ is compatible with $\graph$ if $\prob[\nodes]$ can be consistently factorized into conditional probability distributions for each variable $n \in \nodes$ conditioned on its parents $\Pa[\graph]{n}$:
    \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}] \eq \label{eq:compatibility_as_factorization}\]
    The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ violates \cref{eq:local_markov_property} or equivalently cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is trivially easy to verify whether or not $\prob[\nodes]$ is compatibility with $\graph$; \cref{sec:satisfaction_of_d_sep_relations} describes in detail how this is accomplished.

@@

    Suppose that one were to obtain an inequality on $\graph'$ that contained probabilistic terms such as $\prob[N']$ where $N' \not \in \Inj[\graph]{\graph'}$. At this point, the Inflation lemma is not applicable because it requires probabilistic constraints to consist entirely of distributions $\prob[N']$ where $N'$ is injectable. If for example, $N' \not \in \Inj[\graph]{\graph'}$ could be partitioned into two injectable subsets $N_1' \in \Inj[\graph]{\graph'}$ and $N_2' \in \Inj[\graph]{\graph'}$ that where \term{ancestrally independent} in $\graph'$,
    \[ N_1' \ancestralindep N_2' \iff \An[\graph']{N_1'} \cap \An[\graph']{N_2'}= \emptyset \]
    Then the compatible marginal context $\prob[N']$ must be probabilistically separable $\prob[N'] = \prob[N_1']\prob[N_2']$ due to \cref{eq:local_markov_property}. Since $N_1'$ and $N_2'$ are both injectable, the inflation lemma now applies and the original inequality can be deflated.

>>>>

    Now that the notion of a causal structure $\graph$ is defined along with some standard graph-theoretic notation, it is now possible to precisely define the notion of compatibility between $\graph$ and a probability distribution $\prob[\nodes]$ defined over the nodes of $\graph$. Intuitively, the absence of certain causal connections within $\graph$ should limit the set of possible observations that can be made on $\graph$. Specifically, the causal structure $\graph$ hypothesizes that if each variable $n \in \nodes$ is causally influenced by its parents $\Pa[\graph]{n}$, then each $n$ should be conditionally independent of all other variables:
    \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}] \eq \label{eq:compatibility_as_factorization}\]
    % , except its own descendants when conditioned over $\Pa[\graph]{n}$. This is known as the \term{local Markov property}:\footnote{Here $A \indep B \mid C$ denotes statistical independence between $A$ and $B$ conditional on $C$ where $A,B,C \subseteq \nodes$, i.e. for a given probability distribution $\prob[\nodes]$, $\prob[AB \mid C] = \prob[A \mid C]\prob[B \mid C]$.}
    % \[ n \indep \br{\nodes \setminus \De[\graph]{n}} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is trivially easy to verify whether or not $\prob[\nodes]$ is compatibility with $\graph$; \cref{sec:satisfaction_of_d_sep_relations} describes in detail how this is accomplished.

@@

    Suppose that one were to obtain an inequality on $\graph'$ that contained probabilistic terms such as $\prob[N']$ where $N' \not \in \Inj[\graph]{\graph'}$. At this point, the Inflation lemma is not applicable because it requires probabilistic constraints to consist entirely of distributions $\prob[N']$ where $N'$ is injectable. If for example, $N' \not \in \Inj[\graph]{\graph'}$ could be partitioned into two injectable subsets $N_1' \in \Inj[\graph]{\graph'}$ and $N_2' \in \Inj[\graph]{\graph'}$ that where \term{ancestrally independent} in $\graph'$,
    \[ N_1' \ancestralindep N_2' \iff \An[\graph']{N_1'} \cap \An[\graph']{N_2'}= \emptyset \]
    Then the compatible marginal context $\prob[N']$ must be probabilistically separable $\prob[N'] = \prob[N_1']\prob[N_2']$ due to \cref{eq:compatibility_as_factorization}. Since $N_1'$ and $N_2'$ are both injectable, the inflation lemma now applies and the original inequality can be deflated.
